---
title: "Malaria Health Risk Indicators Dashboard"
output: 
  flexdashboard::flex_dashboard:
    vertical_layout: scroll
runtime: shiny

---

```{r setup, include=FALSE}
install.packages("flexdashboard")
install.packages("plotly")
install.packages("dplyr")

library(flexdashboard)
library(plotly)
library(dplyr)
```

```{r}
recent_year <-read.csv('recent_year_viz.csv')
age_country <-  read.csv('deaths_age_country_viz.csv')
death_country <-  read.csv("deaths_country_viz.csv")

```
Overview in `r max(recent_year$year)` {data-orientation=rows}
=====================================  
This dashboard is split into 2 pages, this dashboard gives an overview of the global incidence rate and death rate of Malaria in the year of `r max(recent_year$year)`. Click on 'Trend over the years' tab to view the other dashboard which gives insight on the trend over a longer horizon.

Row
-------------------------------------

### Average Global Incidence Rate per 1,000 population {.value-box}

```{r}
total_ir <- mean(recent_year$incidence_rate)

# Put valueBox() inside renderValueBox({}) and round off value to 2 decimal place
renderValueBox({
  valueBox(value = round(x=total_ir, digits=2) , icon = "fas fa-users")
})
```


### Average Global Age-Standardised Death Rate per 100,000 people {.value-box}

```{r}

total_dr <- mean(recent_year$death_rate)

# Put valueBox() inside renderValueBox({}) and round off value to 2 decimal place
renderValueBox({
  valueBox(value = round(x=total_dr, digits=2), icon = "fas fa-users")
})
```    

### Number of Deaths in millions {.value-box}

```{r}
total <- sum(age_country$deaths[age_country$year== max(recent_year$year)])/1000000

# Put valueBox() inside renderValueBox({}) and round off value to 2 decimal place
renderValueBox({
  valueBox(value = round(x=total, digits=2), icon = "fas fa-users")
})
```    


Row
-------------------------------------
### Geographical Representation of the Incidence  Rate of Malaria in `r max(recent_year$year)`
```{r}

# light grey boundaries
l <- list(color = toRGB("grey"), width = 0.5)

# specify map projection/options
g <- list(
  showframe = FALSE,
  showcoastlines = FALSE,
  projection = list(type = 'Mercator')
)

fig <- plot_geo(recent_year)
fig <- fig %>% add_trace(
    z = ~incidence_rate, color = ~incidence_rate, colors = 'Reds',
    text = ~entity, locations = ~code, marker = list(line = l)
  )
fig <- fig %>% colorbar(title = 'Incidence Rate per 1,000')
fig <- fig %>% layout(
    geo = g
  )

fig

```
> Malaria is most prevalent in `r recent_year$entity[recent_year$incidence_rate==max(recent_year$incidence_rate)]`.

Row
-------------------------------------
 
### Top 15 Countries with the highest death rates in `r max(recent_year$year)`.
```{r}
topdc <- recent_year %>%
          filter(rank(desc(death_rate))<=15)
```


```{r}
fig2 <- plot_ly(topdc, x = ~death_rate, y = ~entity, color= ~region, type = 'bar', orientation = 'h') %>%
          layout(xaxis = list(title = 'Age-standardised death rate per 100,000 people'), yaxis = list(title = 'Countries'))
fig2
``` 

> The age-standardised death rate (per 100,000 people) is highest amongst the following countries. The top 15 countries are in the `r unique(topdc$region)` region. Click on the Trend over Time tab to view the death rates of these countries over the years.

### Percentage of deaths by age group
In `r max(recent_year$year)`, the percentage of total deaths by age group is:
```{r}
df2 <- age_country[age_country$year == max(recent_year$year),]

fig <- plot_ly(df2, type='pie', labels=~age_group, values=~deaths, 
               textinfo='percent')
fig
```


Trend over the years {data-orientation=rows}
=====================================     
In this dashboard, you can single-click on the legend to remove the item out from the visualisation. Double-clicking an item in the legend will result in the visualisation to highlight that item only.To undo, repeat the same steps.

Row {data-height=500}
-------------------------------------

### Number of deaths by Regions over time
```{r}
max_year_df <- age_country[age_country$year == max(age_country$year),]
min_year_df <- age_country[age_country$year == min(age_country$year),]
```

```{r}

fig3 <- plot_ly(age_country, x = ~year, y = ~deaths, color = ~region, type='bar') %>% 
          layout(barmode = 'stack') %>%
    layout(
      margin = list(b = 50, l = 50) # to fully display the x and y axis labels
    )
fig3
```
> Global deaths has significantly decreased from `r round(x=sum(min_year_df$deaths), digit=2)/1000000` million in the year `r min(age_country$year)` to `r round(x=sum(max_year_df$deaths), digit=2)/1000000` million in the year `r max(age_country$year)`. In the year `r max(age_country$year)`, the `r max_year_df$region[max_year_df$deaths==max(max_year_df$deaths)]` region is the largest contributor to the global deaths. 


### Number of deaths by Age Group over time
```{r}
fig4 <- plot_ly(age_country, x = ~year, y = ~deaths, color = ~age_group, type='bar') %>% 
          layout(barmode = 'stack')

fig4
```   

```{r}
age_df <- age_country[age_country$year == max(age_country$year),]
``` 
> In the year  `r max(age_country$year)`, the age-group that is the most vulnerable to Malaria are those aged `r age_df$age_group[age_df$deaths==max(age_df$deaths)]`. 


Row {data-height=400}
-------------------------------------

### Age-standardised death rates of the top 15 countries in `r max(recent_year$year)` over the years

```{r}

top15 <- unique(topdc$entity)

top15_df <- subset(death_country, entity %in% top15)
  
fig5 <- plot_ly(top15_df,x=~year, y=~death_rate, color=~entity, type='scatter', mode='line') %>%
          layout(xaxis = list(title = 'year'), yaxis = list(title = 'Age-standardised death rate per 100,000 people'))

fig5
```   

